---
title: "Progression Report"
format: gfm
editor: visual
editor_options: 
  chunk_output_type: console
---

This file is used to visually represent the current status of coding data. Included are the reliability scores for coders and topics.

## Data

Raw data can be found in the [data folder](https://github.com/vanatteveldt/issuepositions/tree/main/data) folder and code used to generate statistics and plots is available in [topic_status_report](https://github.com/vanatteveldt/issuepositions/blob/main/src/data-processing/topic_status_report.R) file.

```{r}
#| messages: false
#| warnings: false
#| output: false

library(tidyverse)
library(irr)
library(ggplot2)
library(readr)
library(kableExtra)
library(simplermarkdown)
library(knitr)
library(dplyr)
library(purrr)
library(here)

```

```{r}
#| echo: false
#| messages: false
#| warnings: false
#| output: false


list_units <- function(annotations) {
  units <- read_csv(here::here("data/intermediate/units_tk2023.csv"), 
                    col_select=c("unit_id", "before", "text_hl", "after"), 
                    col_types="cccc") 
  mode <- function(x) names(which.max(table(x)))
  units <- annotations |> 
    left_join(units, by="unit_id") |>
    mutate(text=str_c(str_replace_na(before, ""), text_hl, str_replace_na(after, ""))) |>
    select(-variable, -value, -before, -text_hl, -after) |>
    group_by(unit_id, topic) |>
    mutate(
      jobids = str_c(unique(jobid), collapse=","),
      majority = mode(stance),
      ncoded=n(),
      agreement = mean(stance == majority)) |>
    select(-jobid, -jobtype) |>
    ungroup() |>
    pivot_wider(names_from=coder, values_from=stance)
  if ("NA" %in% colnames(units)) units <- select(units, -"NA") #somehow a column named NA is created with a single value
  units
}

```

```{r}
#| echo: true
#| messages: false
#| warnings: false
#| output: false
gpt_units <- read_csv(here::here("data/intermediate/coded_units_gpt_nshot.csv"))|>
  select(-jobids)

stances <- read_csv(here::here("data/intermediate/stances.csv"))

all_units <- stances |>
  filter(jobtype=="normal") |>
  ungroup() |>
  list_units() |>
  arrange(unit_id) |>
  select(-jobids) |>
  left_join(gpt_units)

normal_units <- stances |>
  filter(jobtype=="normal") |>
  ungroup() |>
  list_units() |>
  arrange(unit_id) |>
  select(-jobids) |>
  left_join(gpt_units)

gpt_issues_all <- read_csv(here::here("data/intermediate/gpt_issues_all.csv")) |>
  filter(logprob >= -5)

```

```{r}
#| echo: false
#| messages: false
#| warnings: false
#| output: false

# CODERS <- c("AvH","MR","WA","JF","NR","S", "ING","NK","SH","OY","KN","NPR","JE","AM", "GPT-1shot", "GPT-5shot", "GPT-16shot", "GPT-35shot")
CODERS <- c("NR","NK","SH","KN","NPR")

# # Identify the abbreviations of coders that are present in all_units
present_coders <- intersect(CODERS, names(all_units))

# # Convert only the columns for the present coders to numeric values
all_units_numeric <- all_units |> 
  mutate(across(all_of(present_coders), ~ as.numeric(factor(., levels=c("L", "N", "R")))))
normal_units_numeric <- normal_units |> 
  mutate(across(all_of(present_coders), ~ as.numeric(factor(., levels=c("L", "N", "R")))))

# Overall Reliability calculation
alpha <- function(all_units_numeric) {
  select(all_units_numeric, all_of(present_coders)) |>
    # Prepare numeric data as a matrix for Krippendorff calculation
    as.matrix() |>
    t() |>
    irr::kripp.alpha(method="nominal")
}

# Pairwise reliability
pairwise_alpha_new <- function(all_units_numeric) {
  coders <- present_coders
  result <- tibble(coder1 = character(), coder2 = character(), alpha = numeric())
  
  for (i in 1:(length(coders) - 1)) {
    for (j in (i + 1):length(coders)) {
      coder1 <- coders[i]
      coder2 <- coders[j]
      
      
      # Filter data to include only coder columns
      sub_data <- all_units_numeric |>
        select(all_of(c(coder1, coder2))) |>
        na.omit()
      
      # Check if both coder columns are available
      if (sum(colnames(sub_data) %in% c(coder1, coder2)) == 2 && nrow(sub_data) > 1) {
        tryCatch({
          alpha_value <- irr::kripp.alpha(t(as.matrix(sub_data)), method = "nominal")
          result <- result |>
            bind_rows(tibble(coder1 = coder1, coder2 = coder2, alpha = alpha_value$value))
        }, error = function(e) {
          message("Error calculating alpha for ", coder1, " and ", coder2, ": ", e$message)
        })
      } else {
        #message("Insufficient data for calculating alpha between ", coder1, " and ", coder2)
      }
          }
        }
        
  return(result)
}

topic_report  <- all_units  |> group_by(topic) |> summarize(n=n(), ncoded=mean(ncoded), agreement=mean(agreement))

topic_alphas <- topic_report |> pull(topic) |> unique() |> map(function(t) tibble(topic=t, alpha=filter(normal_units_numeric, topic==t) |> alpha() |> pluck("value"))) |> list_rbind()
left_join(topic_report, topic_alphas) |> arrange(-n)

table(all_units$majority[all_units$topic == "Government"])

overall_kripp_alpha <- alpha(all_units_numeric)
normal_kripp_alpha <- alpha(normal_units_numeric)

pairwise_kripp_alpha <- pairwise_alpha_new(normal_units_numeric)


#plotting reliability values between coders

plot_pairwise_kripp_alpha <- function(pairwise_kripp_alpha) {
  # Convert coder columns to factors to ensure correct plotting order
  pairwise_kripp_alpha <- pairwise_kripp_alpha %>%
    mutate(coder1 = factor(coder1, levels = unique(c(coder1, coder2))),
           coder2 = factor(coder2, levels = unique(c(coder1, coder2))))
  
  # Plot heatmap
  ggplot(pairwise_kripp_alpha, aes(x = coder1, y = coder2, fill = alpha, label=round(alpha,2))) +
    geom_tile(color = "white") +
    geom_text(color = "white") +
    scale_fill_gradient2(low = "darkred", high = "darkgreen", mid = "gold", midpoint = 0.5,
                         name = "Krippendorff's Alpha") +
    labs(title = "Pairwise Krippendorff's Alpha Between Coders",
         x = "",
         y = "") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
    coord_fixed()
}

# create topic info table


topic_alpha <- function(df, topic_name) {
  all_units_numeric <- mutate(df, across(all_of(present_coders), ~ as.numeric(factor(.))))
  result <- all_units_numeric[all_units_numeric$topic == topic_name, ] |>
  select(all_of(present_coders)) |>
    # Prepare numeric data as a matrix for Krippendorff calculation
    as.matrix() |>
    t() |>
    irr::kripp.alpha(method="nominal") 
  return(round(result$value,2))
}


```

## Coder Reliability

The overall reliability across is a Krippendorff's alpha of **`r round(normal_kripp_alpha$value, 2)`** (excluding test and no-agreement units)

```{r}
#| echo: false
#| messages: false
#| warnings: false
#| output: true
#| label: plot-alpha

# #| fig.cap: "Reliability scores between coders"

plot_pairwise_kripp_alpha(pairwise_kripp_alpha)

```

## Topic Reliability

The following table reports the progression of issues coded and the current reliability (calculated using Krippendorff's alpha) for each topic.

<center>

```{r results='asis', echo=FALSE, message=F}

topic_status <- function(df, total_df){
  topic_info <- tibble(topic_name = character(), completed_count = numeric(), total_count = numeric(), percentage_done = character(), reliability_α = numeric())  # Initialize as character to hold the formatted percentage
  
  total_counts <- total_df |> 
    group_by(topic) |> 
    summarize(total_count=length(unique(unit_id))) |> 
    na.omit()
  completed_count <- df |> 
    group_by(topic) |> 
    summarize(completed_count=length(unique(unit_id))) |> 
    na.omit()
  
  alphas <- completed_count |> 
    pull(topic) |> 
    unique() |>
    
    map(function(topic) tibble(topic=topic, alpha=topic_alpha(df, topic)), .progress = T) |>
    list_rbind()
  
  left_join(total_counts, completed_count) |>
    replace_na(list(completed_count=0)) |>
    mutate(percentage_done = str_c(round(100*completed_count/total_count), "%")) |>
    left_join(alphas) 
}


topic_info <- topic_status(all_units, gpt_issues_all)

# Define custom column names
custom_headers <- c("Topic", "Completed", "Total", "Percentage Done", "Reliability (α)")


# Simple table for GFM
# Use kable for Markdown output suitable for GFM

kable(topic_info, format = "markdown", col.names = custom_headers, align = 'c')
```

</center>

```{r}
#| echo: false
#| messages: true
#| warnings: false
#| label: pairwise-plots
#| results: asis


for  (topic_name in sort(unique(all_units$topic))) {
  cat("\n## Topic:", topic_name, "\n\n")
  
  topic_alphas <- all_units_numeric |>
    filter(topic == topic_name) |>
    pairwise_alpha_new()
  
  plot <- plot_pairwise_kripp_alpha(topic_alphas)
  print(plot)
  
  a <- alpha(filter(all_units_numeric, topic == topic_name))
  
  cat("\n\n#### Topic α score:", a$value, "\n\n")
  
  cat('\n\n')
}


```
